A Simple Type Theory for Category Theory

### 2022-02-11 paper/lipics-v2021-sample-article.tex
abstract update

%TODO mandatory: add short abstract of the document
\begin{abstract}
  We present a domain-specific type theory for formal category theory
  constructions and proofs. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to a form of ordred
  logic, we guarantee that all mappings between categories are
  functorial and all transformations are natural by construction, with
  no separate proof necessary. We demonstrate the utility of the type
  theory by proving a form of the Yoneda lemma which reads as the
  classical proof, but with all naturality side-conditions
  automatically satisfied.

  We show that the type theory is sound and complete for virtual
  equipments, a kind of double category which is known to model not
  just ordinary categories but enriched and internal categories as
  well. This means that proofs and constructions in our theory hold
  not just for ordinary categories but also for a broad class of
  generalized categories. We demonstrate that the type theory
  simplifies working with virtual equipments compared to their
  algebraic description in the literature.
\end{abstract}

>>>>

%TODO mandatory: add short abstract of the document
\begin{abstract}
  We present a domain-specific type theory for constructions and
  proofs in category theory. The type theory axiomatizes notions of
  category, functor, profunctor and a generalized form of natural
  transformations. By restricting the type theory to a form of ordered
  logic, we guarantee that all mappings between categories are
  functorial and all transformations are natural by construction, with
  no separate proof necessary. Important category theoretic proofs
  such as the Yoneda lemma and Co-yoneda lemma become simple type
  theoretic proofs about the relationship between unit, tensor and
  (ordered) function types.

  We show that the type theory is sound for models called
  \emph{virtual equipments}, a kind of double category which is known
  to model not just ordinary categories but enriched, internal and
  generalized multi-categories as well. This means that while proofs
  and constructions in our theory read like typical category-theoretic
  constructions, the proofs automatically apply to these generalized
  categories. Our type theory simplifies arguments over the algebraic
  description of virtual equipments by modeling the algebraic
  operation of restriction of profunctors along functors as instead
  the strict syntactic substitution.
\end{abstract}
